All Graphs Lead to Rome: Learning Geometric and Cycle Consistent Representations with Graph CNNs

### 2018-11-16 cvpr2019/main.tex
Added some more text and updated figure

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
% Include other packages here, before hyperref.
\usepackage[outdir=./]{epstopdf}

@@

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

@@

\section{Method}
Our goal is to learn optimal features that capture multiple image views by way of filtering out noisy feature matches.
The input to our algorithm is a set of features and noisy correspondences, and the output is a new set of features where the pairwise similarities of these features correspond to the true matches.
We do this by learning training the new set of feature embeddings to be cycle consistent.
We formulate this problem in terms of the correspondence graph of the features.
Vectors $\mat{x}$ and matrices $\mat{A}$ are denoted with boldface, and the $i^{th}$ row of a matrix is denoted $(\mat{A})_i$.
\subsection{Correspondence Graph}
We assume we have an initial set of feature matches represented as a graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$.
Each vertex $v$ of the graph is a feature with its associated descriptor $f_v$. 
The graph is represented by its Weighted Adjacency Matrix $\mat{A}(\mathcal{G}) \in \bR^{n \times n}$, where $(\mat{A}(\mathcal{G}))_{ij}$ is the strength of the match between nodes $i$ and $j$. We also have the positive diagonal degree matrix $\mat{D}(\mathcal{G}) \in \bR^{n \times n}$, with $\mat{D}(\mathcal{G})_{ii} = \sum_j (\mat{A}(\mathcal{G}))_{ij}$. For brevity we denote these matrices $\mat{A}$ and $\mat{D}$ respectively.
We can get our an embedding matrix $\mat{E}_0$ by concatenating the features $f_v$ from each of the vertices.
We use $\mat{E}_0$ as the initializiation for our learning algorithm.
We add any additional knowledge we have into the embedding as well (e.g. scale, orientation, etc.).
Putative correspondences are computed from these features and represented by weighted edges in which the weight gives the strength of the match.
While there are many interesting methods for computing these putative correspondences \cite{suh2015subgraph, yi2018learning}, we do not explore them in this work.

@@

\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.9\linewidth]{figures/UniverseOfFeatures.pdf}
\end{center}
  \caption{
    Illustration of the universe of features.
    Each feature in each image corresponds to a 3D point in the scene.
    Therefore, we can get the points to have a cycle consistent embedding by mapping each feature in each image to the one-hot vector of its corresponding 3D point.
    While there can be many features, there are much fewer 3D points and thus this corrsponds to a low rank factorization of the corrspondence matrix.
    Best viewed in color.
  }
\label{fig:universefeatures}
\label{fig:onecol}
\end{figure}

@@

@@

@@

\begin{figure*}
\begin{center}
  % \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{figures/TrainingCurvesFull.pdf}
  \end{center}
    \caption{
      Training curves with and without Geometric Training loss.
      The geometric training loss improves testing performance.
      Shown as horizontal lines are the state of the art optimization based baselines.
      Even with a simple network we compare well with them.
    }
  \label{fig:short}
\end{figure*}

\begin{figure*}
\begin{center}
  % \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{figures/experiments01-eps-converted-to.pdf}
  \end{center}
     \caption{Experimental results from Rome16K}
  \label{fig:short}
\end{figure*}

\hline
\end{tabular}
\end{center}
\caption{
Results on Synthetic correspondence graphs.
The `Same Point Similarities' column is the mean and standard deviation of similarities for true corresponding points, while the `Different Point Similarities' is the same for points that do not correspond.
For the `Same Point Similarities' column higher is better, and for `Different Point Similarities' lower is better.
Losses tested against ground truth correspondence graph adjacency matrices.
Our method was not trained on ground truth correspondences but using unsupervised methods.
}
\label{fig:synthtable}
\end{table*}
\subsection{Synthetic Graph Dataset}
We first test our method on synthetically generated data as a simple proof of concept.
To generate the data, we generate $p$ points, each with its own randomly generated descriptor.
To create the graph, we generate random permutation matrices, with a noise applied to it after generated.
The initial descriptors we use as input we created using the true descriptor plus some added Gaussian noise.
No geometric losses were added during training for these experiments.
However, the method was robust in testing with different noise functions and parameters.
The normalized noisy input descriptors are our baseline - they correlate with the true values but do not perserve the structure well.
However, with the GNN we are able to recover the true structure very well, as shown in \ref{fig:synthtable}.
All experiments were run with a 12 layer GCN with the ReLU nonlinearity and skip connections.
All were trained with the Adam optimizer \cite{kingma2014adam} and a learning rate $10^{-4}$

\subsection{Rome 16K Graph Dataset}
We  use the Rome16K dataset \cite{li2010location} to test our algorithm in real world settings.
Rome16K consistes of 16 thousand images of various historical sites in Rome extracted from Flickr, along with the 3D structure of the sites provided by bundle adjustment.
While not a standard dataset to test cycle consistency, other datasets had insufficient data to train a network on.
Rome16K is typically used to test bundle adjustment methods.
Therefore, to use for our method, we extract image triplets and quadtruplets with overlap of 80 points or more to test our algorithm on, with the points established as corresponding in the given bundle adjument output.
For the initial embedding we use the original 128 dimensional SIFT descriptors, normalized to have $L_2$ norm 1, the calibrated x-y position, the orientation, and log scale of the SIFT feature.
% For the outlier datasets, we take 79 correspondences and 1 non-corresponding point in each image.

For these experiments we train with the $L1$ norm, and ablate the geometric side losses.
We evaluate on a test set using the ground truth adjacency matrix which we compute from the bundle adjustment given by the Rome16K dataset.
Traditional methods of evaluation use Precision and Recall, after making the labels 0 or 1.
We forgo this as our method was trained to output soft labels - we in turn make the labels soft for the optimization based methods.
For this method to work, we need the dimension of the embedding to be at least the number of unique points in the scene.
Picking the correct number is difficult a-priori, and is a problem with all cycle consistency based methods.
Here we use the ground truth dimension of the embedding to test both our method and the baselines.

\begin{table*}
\begin{center}
\begin{tabular}{|l|c|c|c|}
\hline
Method                                                    & $L_1$ Loss        & $L_2$ Loss        & BCE Loss \\

\hline\hline\hline
MatchALS \cite{zhou2015multi} 15 Iterations               & 0.052 $\pm$ 0.003 & 0.010 $\pm$ 0.002 & 0.107 $\pm$ 0.038 \\ \hline
MatchALS \cite{zhou2015multi} 100 Iterations              & 0.007 $\pm$ 0.003 & 0.007 $\pm$ 0.003 & 0.220 $\pm$ 0.099 \\ \hline
% MatchALS \cite{zhou2015multi} 400 Iterations              & 0.007 $\pm$ 0.003 & 0.007 $\pm$ 0.003 & 0.225 $\pm$ 0.104 \\ \hline
PGDDS \cite{leonardos2016distributed} 15 Iterations       & 0.016 $\pm$ 0.002 & 0.006 $\pm$ 0.002 & 0.045 $\pm$ 0.037 \\ \hline
PGDDS \cite{leonardos2016distributed} 100 Iterations      & 0.013 $\pm$ 0.002 & 0.005 $\pm$ 0.002 & 0.039 $\pm$ 0.038 \\ \hline
% PGDDS \cite{leonardos2016distributed} 200 Iterations      & 0.013 $\pm$ 0.002 & 0.005 $\pm$ 0.002 & 0.039 $\pm$ 0.039 \\ \hline
Spectral                                                  & 0.054 $\pm$ 0.005 & 0.018 $\pm$ 0.004 & 0.114 $\pm$ 0.028 \\ \hline \hline
GNN, 12 Layers (ours)                                     & 0.027 $\pm$ 0.003 & 0.017 $\pm$ 0.003 & 0.701 $\pm$ 0.002 \\
\hline
\end{tabular}
\end{center}
% save-rome16kgeom0-normedskip2-loss-l1-geom-10-0 (028249) -> train : +3.1467e-02, test : +3.8864e-02, gt_l1 : +2.6321e-02, gt_l2 : +2.6321e-02

\caption{
Results on Rome16K Correspondence graphs, showing the mean and standard deviation of the $L_1$ and $L_2$.
Our method was not trained on ground truth corresopndences but using unsupervised methods and geometric side losses.
As our method gives soft labels, we use cannot use precision or recall as is standard in testing cycle consistency \cite{zhou2015multi}.
Thus we test against ground truth correspondence graph adjacency matrices gotten from the bundle adjustment output.
}
\end{table*}

@@

We have shown a novel method for training feature matching.
We have demonstrated its effectiveness using experiments on the Rome16K dataset.
This method can be employed even when labelled training data is not available, but can still incorporate side geometric information.
For future work we will investigate using more robust losses for outlier rejection and using higher order geometric constraints, such as the trifocal tensor, as additional loss terms.
A long term goal is to incorporate learning image level features into this pipeline
We hope to make this method distributed - with group norm, we cannot distribute in its current form.

>>>>

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
% Include other packages here, before hyperref.
% \usepackage[outdir=./]{epstopdf}

@@

\def\cvprPaperID{3510} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

@@

\section{Method}

\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.9\linewidth]{figures/UniverseOfFeatures.pdf}
\end{center}
  \caption{
    Illustration of the universe of features.
    Each feature in each image corresponds to a 3D point in the scene.
    Therefore, we can get the points to have a cycle consistent embedding by mapping each feature in each image to the one-hot vector of its corresponding 3D point.
    While there can be many features, there are much fewer 3D points and thus this corrsponds to a low rank factorization of the corrspondence matrix.
    Best viewed in color.
  }
\label{fig:universefeatures}
\label{fig:onecol}
\end{figure}
Our goal is to learn optimal features that capture multiple image views by way of filtering out noisy feature matches.
The input to our algorithm is a set of features and noisy correspondences, and the output is a new set of features where the pairwise similarities of these features correspond to the true matches.
We do this by learning training the new set of feature embeddings to be cycle consistent.
We formulate this problem in terms of the correspondence graph of the features.
Vectors $\mat{x}$ and matrices $\mat{A}$ are denoted with boldface, and the $i^{th}$ row of a matrix is denoted $(\mat{A})_i$.
\subsection{Correspondence Graph}
We assume we have an initial set of feature matches represented as a graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$.
Each vertex $v$ of the graph is a feature with its associated descriptor $f_v$. 
The graph is represented by its Weighted Adjacency Matrix $\mat{A}(\mathcal{G}) \in \bR^{n \times n}$, where $(\mat{A}(\mathcal{G}))_{ij}$ is the strength of the match between nodes $i$ and $j$. We also have the positive diagonal degree matrix $\mat{D}(\mathcal{G}) \in \bR^{n \times n}$, with $\mat{D}(\mathcal{G})_{ii} = \sum_j (\mat{A}(\mathcal{G}))_{ij}$. For brevity we denote these matrices $\mat{A}$ and $\mat{D}$ respectively.
We can get our an embedding matrix $\mat{E}_0$ by concatenating the features $f_v$ from each of the vertices.
We use $\mat{E}_0$ as the initializiation for our learning algorithm.
We add any additional knowledge we have into the embedding as well (e.g. scale, orientation, etc.).
Putative correspondences are computed from these features and represented by weighted edges in which the weight gives the strength of the match.
While there are many interesting methods for computing these putative correspondences \cite{suh2015subgraph, yi2018learning}, we do not explore them in this work.

@@

@@

\hline
\end{tabular}
\end{center}
\caption{
Results on Synthetic correspondence graphs.
The `Same Point Similarities' column is the mean and standard deviation of similarities for true corresponding points, while the `Different Point Similarities' is the same for points that do not correspond.
For the `Same Point Similarities' column higher is better, and for `Different Point Similarities' lower is better.
Losses tested against ground truth correspondence graph adjacency matrices.
Our method was not trained on ground truth correspondences but using unsupervised methods.
}
\label{fig:synthtable}
\end{table*}

@@

\begin{table*}
\begin{center}
\begin{tabular}{|l|c|c|c|}
\hline
Method                                                    & $L_1$ Loss        & $L_2$ Loss         \\

\hline\hline\hline
MatchALS \cite{zhou2015multi} 15 Iterations               & 0.052 $\pm$ 0.003 & 0.010 $\pm$ 0.002 \\ \hline
MatchALS \cite{zhou2015multi} 25 Iterations               & 0.045 $\pm$ 0.007 & 0.009 $\pm$ 0.003 \\ \hline
MatchALS \cite{zhou2015multi} 50 Iterations               & 0.016 $\pm$ 0.008 & 0.007 $\pm$ 0.003 \\ \hline
MatchALS \cite{zhou2015multi} 100 Iterations              & 0.007 $\pm$ 0.003 & 0.007 $\pm$ 0.003 \\ \hline
% MatchALS \cite{zhou2015multi} 400 Iterations              & 0.007 $\pm$ 0.003 & 0.007 $\pm$ 0.003 \\ \hline
PGDDS \cite{leonardos2016distributed} 15 Iterations       & 0.016 $\pm$ 0.002 & 0.006 $\pm$ 0.002 \\ \hline
PGDDS \cite{leonardos2016distributed} 50 Iterations       & 0.013 $\pm$ 0.002 & 0.005 $\pm$ 0.002 \\ \hline
PGDDS \cite{leonardos2016distributed} 100 Iterations      & 0.013 $\pm$ 0.002 & 0.005 $\pm$ 0.002 \\ \hline
% PGDDS \cite{leonardos2016distributed} 200 Iterations      & 0.013 $\pm$ 0.002 & 0.005 $\pm$ 0.002 & 0.039 $\pm$ 0.039 \\ \hline
Spectral                                                  & 0.054 $\pm$ 0.005 & 0.018 $\pm$ 0.004 \\ \hline \hline
GNN, 12 Layers (ours)                                     & 0.025 $\pm$ 0.003 & 0.016 $\pm$ 0.003 \\
\hline
\end{tabular}
\end{center}
% save-rome16kgeom0-normedskip2-loss-l1-geom-10-0 (028249) -> train : +3.1467e-02, test : +3.8864e-02, gt_l1 : +2.6321e-02, gt_l2 : +2.6321e-02

\caption{
Results on Rome16K Correspondence graphs, showing the mean and standard deviation of the $L_1$ and $L_2$.
Our method was not trained on ground truth corresopndences but using unsupervised methods and geometric side losses.
As our method gives soft labels, we use cannot use precision or recall as is standard in testing cycle consistency \cite{zhou2015multi}.
Thus we test against ground truth correspondence graph adjacency matrices computed from the bundle adjustment output.
Our method performs better than 25 iteration of an optimizaiton based method, but does not perform as well as 50 iterations.
Our network has only 12 layers, thus we have much greater efficiency per iteration.
}
\end{table*}

@@

\begin{figure*}
\begin{center}
  % \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{figures/TrainingCurves.pdf}
  \end{center}
    \caption{
      Training curves with and without Geometric Training loss.
      The geometric training loss improves testing performance.
      Shown as horizontal lines are the state of the art optimization based baselines.
      Even with a simple network we compare well with them.
    }
  \label{fig:short}
\end{figure*}

\begin{figure*}
\begin{center}
  % \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{figures/ExampleOutput.pdf}
  \end{center}
     \caption{Example output of our network. (a) Similarity Matrix of the of the embeddings (b) Histogram of feature similarities for pairs which correspond to the same 3D point, comparing our output features with the original SIFT features (c) Histogram of feature similarities for pairs which correspond to different 3D points, comparing our output features with the original SIFT features}
  \label{fig:short}
\end{figure*}
\subsection{Synthetic Graph Dataset}
We first test our method on synthetically generated data as a simple proof of concept.
To generate the data, we generate $p$ points, each with its own randomly generated descriptor.
To create the graph, we generate random permutation matrices, with a noise applied to it after generated.
The initial descriptors we use as input we created using the true descriptor plus some added Gaussian noise.
No geometric losses were added during training for these experiments.
However, the method was robust in testing with different noise functions and parameters.
The normalized noisy input descriptors are our baseline - they correlate with the true values but do not perserve the structure well.
However, with the GNN we are able to recover the true structure very well, as shown in \ref{fig:synthtable}.
All experiments were run with a 12 layer GCN with the ReLU nonlinearity and skip connections.
All were trained with the Adam optimizer \cite{kingma2014adam} and a learning rate $10^{-4}$

\subsection{Rome 16K Graph Dataset}
We  use the Rome16K dataset \cite{li2010location} to test our algorithm in real world settings.
Rome16K consistes of 16 thousand images of various historical sites in Rome extracted from Flickr, along with the 3D structure of the sites provided by bundle adjustment.
While not a standard dataset to test cycle consistency, other datasets had insufficient data to train a network on.
Rome16K is typically used to test bundle adjustment methods.
Therefore, to use for our method, we extract image triplets and quadtruplets with overlap of 80 points or more to test our algorithm on, with the points established as corresponding in the given bundle adjument output.
For the initial embedding we use the original 128 dimensional SIFT descriptors, normalized to have unit $L_2$ norm, the calibrated x-y position, the orientation, and log scale of the SIFT feature.
% For the outlier datasets, we take 79 correspondences and 1 non-corresponding point in each image.

For these experiments we train with the $L1$ norm and geometric consistency losses.
We evaluate on a test set using the ground truth adjacency matrix which we compute from the bundle adjustment given by the Rome16K dataset.
Traditional methods of evaluation use Precision and Recall, after making the labels 0 or 1.
We forgo this as our method was trained to output soft labels - we in turn make the labels soft for the optimization based methods.
For this method to work, we need the dimension of the embedding to be at least the number of unique points in the scene.
Picking the correct number is difficult a-priori, and is a problem with all cycle consistency based methods.
Here we use the ground truth dimension of the embedding to test both our method and the baselines.
For implementation details, our network has 12 layers, and was trained with the Adam optimizer \cite{kingma2014adam} with a learning rate of $10^{-4}$, with an exponentially decaying learning rate. We incorporate skip connections between the input, $6^{th}$, and $12^{th}$ layers, which greatly improved training convergence and testing performance (see supplemental material).

We compare our method to spectral and optimization based baselines with different maximum iteration cutoffs. Our method is siginificatnly better than a simple spectral based optimization. Furthermore, our network, which only has 12 layers has the accuracy of the optimization based method MatchALS \cite{zhou2015multi} between 25 and 50 iterations. This inidicates that our method learns a more efficient update with each layer compared with an iteration in the optimization based methods.

@@

We have shown a novel method for training feature matching.
This method can be employed even when labelled training data is not available, but can still incorporate geometric consistency into its losses.
We have demonstrated its effectiveness using experiments on the Rome16K dataset compared to the traditional optimization based baselines.
For future work we will investigate using more robust losses for outlier rejection and using higher order geometric constraints, such as the trifocal tensor, as additional loss terms.
A long term goal is to incorporate learning image level features into this pipeline
Moreover we hope to make this method distributed - with group norm, we cannot distribute in its current form.
